Highly Available Transactions: Virtues and Limitations

### 2013-03-31 writeup/hat-analysis.tex

\begin{table}[t!]
\begin{tabular}{| c | p{6cm} | }\hline
HA & Transactional Atomicity (TA), Read Uncommitted (RU), Read
Committed (RC), Item Cut Isolation (P-CI), Predicate Cut Isolation
(P-CI), Monotonic Reads (MR), Monotonic Writes (MW), Writes Follow
Reads (WFR)\\\hline Sticky-HA & Read Your Writes (RYW), PRAM,
Causal\\\hline Unavailable & Cursor Stability (CS)\lostupdate,
Snapshot Isolation (SI)\lostupdate, Repeatable Read
(RR)\lostupdate\rwskew, One-Copy Serializability
(1SR)\lostupdate\rwskew, Recency\linearizable, Safe\linearizable,
Regular\linearizable, Strong 1SR\lostupdate\rwskew\linearizable
\\\hline
\end{tabular}
\caption{Summary of highly available, sticky highly available, and
  unavailable models considered in this paper. Unavailable models are
  labeled by cause of unavailability: preventing lost
  update\lostupdate, preventing write skew\rwskew, and requiring
  recency guarantees\linearizable.}
\label{table:hatcompared}
\end{table}

@@

 \node[draw=none,fill=none] (ta) at (-.2, 1.3) {TA};
 \node[draw=none,fill=none] (mr) at (2.4, 0) {MR};
 \node[draw=none,fill=none] (mw) at (3.6, 0) {MW};
 \node[draw=none,fill=none] (wfr) at (6.3,0) {WFR};
 \node at (5,0) [sticky] (ryw) {RYW};

 \node[noha](recency) at (7.7, 0) {recency};
 \node[noha](safe) at (7.7, 1) {safe};
 \node[noha](regular) at (7.7, 2) {regular};
 \node[noha](linearizable) at (7.7, 3) {linearizable};
 \node at (3.6, 2) [sticky] (causal) {causal};
 \node at (3.6, 1) [sticky] (pram) {PRAM};
 \node[noha] (cs) at (-1.2, 1.8) {CS};
 \node[noha] (rr) at (-.2, 2.7) {RR};
 \node[noha] (si) at (2.2, 2.4) {SI};
 \node[noha] (1sr) at (1.2, 3.2) {1SR};
 \node[noha] (ssr) at (3.6, 3.6) {Strong-1SR};

@@

 \draw [->, red] (rc) -- (cs);
 \draw [->, red] (cs) -- (rr);
 \draw [->, red] (pci) -- (si);
 \draw [->, red] (ici) -- (rr);
 \draw [->, red] (rr) -- (1sr);
 \draw [->, red] (si) -- (1sr);
 \draw [->, red] (ta) -- (si);
 \draw [->, red] (ta) -- (rr);
 \draw [->, red] (causal) -- (ssr);

>>>>

\begin{table}[t!]
\begin{tabular}{| c | p{6cm} | }\hline
HA & Transactional Atomicity (TA), Read Uncommitted (RU), Read
Committed (RC), Item Cut Isolation (P-CI), Predicate Cut Isolation
(P-CI), Monotonic Reads (MR), Monotonic Writes (MW), Writes Follow
Reads (WFR)\\\hline Sticky-HA & Read Your Writes (RYW), PRAM,
Causal\\\hline Unavailable & Cursor Stability (CS)\lostupdate,
Snapshot Isolation (SI)\lostupdate, Repeatable Read
(RR)\lostupdate\rwskew, One-Copy Serializability
(1SR)\lostupdate\rwskew, Recency\linearizable, Safe\linearizable,
Regular\linearizable, Linearizability\linearizable, Strong
1SR\lostupdate\rwskew\linearizable \\\hline
\end{tabular}
\caption{Summary of highly available, sticky highly available, and
  unavailable models considered in this paper. Unavailable models are
  labeled by cause of unavailability: preventing lost
  update\lostupdate, preventing write skew\rwskew, and requiring
  recency guarantees\linearizable.}
\label{table:hatcompared}
\end{table}

@@

 \node[draw=none,fill=none] (ta) at (-.2, 1.3) {TA};

 \node[draw=none,fill=none] (mr) at (3.6, 0) {MR};
 \node[draw=none,fill=none] (mw) at (4.8, 0) {MW};
 \node[draw=none,fill=none] (wfr) at (2.4,0) {WFR};
 \node at (6.1,0) [sticky] (ryw) {RYW};

 \node[noha](recency) at (7.7, 0) {recency};
 \node[noha](safe) at (7.7, 1) {safe};
 \node[noha](regular) at (7.7, 2) {regular};
 \node[noha](linearizable) at (7.7, 3) {linearizable};
 \node at (4.8, 2) [sticky] (causal) {causal};
 \node at (4.8, 1) [sticky] (pram) {PRAM};
 \node[noha] (cs) at (-1.2, 1.8) {CS};
 \node[noha] (rr) at (-.2, 2.7) {RR};
 \node[noha] (si) at (2.2, 2.4) {SI};
 \node[noha] (1sr) at (1.2, 3.2) {1SR};
 \node[noha] (ssr) at (3.85, 3.6) {Strong-1SR};

@@

 \draw [->, red] (rc) -- (cs);
 \draw [->, red] (cs) -- (rr);
 \draw [->, red] (pci) -- (si);
 \draw [->, red] (ici) -- (rr);
 \draw [->, red] (rr) -- (1sr);
 \draw [->, red] (si) -- (1sr);
 \draw [->, red] (ta) -- (si);
 \draw [->, red] (ta) -- (rr);
 \draw [->, red] (causal) -- (linearizable);
 \draw [->, red] (ryw) -- (safe);

### 2013-03-31 writeup/hat-vldb.tex
New intro

\documentclass{vldb}
\usepackage{times, graphicx, url, xcolor, multicol, minipage, subfig, tikz, draftwatermarkbottom, draftwatermarktop}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)

@@

\begin{abstract}
To minimize network latency and handle the possibility of partial
failure and network partitions, many modern distributed systems eschew
transactional functionality, which provides strong semantic guarantees
for groups of multiple operations over multiple data items. In this
work, we consider the problem of achieving transactional guarantees
with high availability, or without suffering from unavailability
during system partitions or incurring high network latency.  We
introduce a new taxonomy of high availability and analyze a
combination of ACID and distributed consistency guarantees to identify
which can and cannot be achieved by Highly Available Transaction (HAT)
systems. In doing so, we analytically and experimentally quantify the
benefit of providing high availability as well as the associated
semantic compromises.
\end{abstract}

@@

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
%pagebreak

%\begin{appendix}

%\end{appendix}

\end{document}

>>>>

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax

\usepackage{times, graphicx, url, xcolor, multicol, minipage, subfig, tikz, draftwatermarkbottom, draftwatermarktop, amsthm}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{xspace}

\newcommand{\etal}{{\em et al.}\xspace}
\newcommand{\etc}{{\em etc}\xspace}
\newcommand{\cf}{{\em c.f.,}\xspace}
\newcommand{\versus}{{\em vs.}\xspace}
\newcommand{\ie}{{\em i.e.,}\xspace}
\newcommand{\eg}{{\em e.g.,}\xspace}

\theoremstyle{definition}
\newtheorem{observation}{Observation}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}

@@

\begin{abstract}
To minimize network latency and remain online during partial failure
and network partitions, many modern distributed data storage systems
eschew transactional functionality, which provides strong semantic
guarantees for groups of multiple operations over multiple data
items. In this work, we consider the problem of providing Highly
Available Transactions (HATs), or transactional guarantees that do not
suffer from unavailability during system partitions or incur high
network latency.  We introduce a new taxonomy of high availability and
analyze a existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. In doing so,
we analytically and experimentally quantify the benefit of providing
high availability as well as the associated semantic compromises.
\end{abstract}

In the debate over NoSQL systems versus traditional DBMS technologies for distributed data management, one key issue is that many NoSQL systems aim for availability in the face of partitions, and provide no, or restricted, support for transactions that group multiple operations over multiple data items. The authors recently advocated Highly Available Transaction (HAT) systems, where general transactions are provided with ACID-like semantic guarantees such as weak isolation levels, in ways that remain available during partitions and avoid high network latency. Here we introduce a new taxonomy of partition-availability properties. We analyse existing ACID and distributed consistency guarantees, showing the availability that can be achieved for each.  We analytically and experimentally estimate the performance gains and semantic compromises of providing high availability, compared to traditional serializable transactions in an internet-scale setting.

@@

%APPENDIX is optional.
% ****************** APPENDIX **************************************
% Example of an appendix; typically would start on a new page
\pagebreak

\begin{appendix}
\input{availability-appendix.tex}

\input{hat-appendix.tex}
\end{appendix}
\end{document}
